---
title: "ESM 206 Lab 10"
author: "YOUR NAME HERE"
date: "December 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab 10 Objectives:

- Correlation matrices
- Binary logistic regression
- Dates/times with lubridate
- Heatmap of solar irradiation in Santa Barbara, Hilo (HI), and Fairbanks (AK) in 2010
- Aquaculture with gganimate

# PART 0. Load libraries
```{r packages}

library(tidyverse) # The tidyverse!
library(corrplot) # For correlation matrices
library(janitor) # For cleaning up column names
library(lubridate) # For dealing with dates & times
library(gganimate) # For adding animation to ggplot graphics
library(ggimage) # For updating graph components with images


```

# PART 1. Correlation matrix (World Bank world environmental factors data)

Compiled World Bank data, accessed from: https://www.kaggle.com/zanderventer/environmental-variables-for-world-countries#World_countries_env_vars.csv

- Read in the "world_env_vars.csv file" as env_var
- Find correlations between all variables, columns 2:28
- Plot the correlation matrix (corrplot)

```{r env_corr}

# reading csv
env_var = read_csv("world_env_vars.csv") %>% 
  na.omit

# calculate Pearson's r for correlations from column 2-28
cor_df = cor(env_var[2:28])

# Make a correlation matrix... ellipse shows the dirctionality
# tl.col = text color, tl.cex = text size
corrplot(cor_df, type = "upper", method = "ellipse", tl.col = "navy blue", tl.cex = 0.5)

```

# PART 2. Binary Logistic Regression (Donner Party Data)

Use the 'glm' function for fitting *generalized linear models* (the *logit* - log odds of survival, in our case, will be linearly related to Sex and Age. So we expect the final model to look something like this: 

$$Log Odds (Survival) = \beta_0 + \beta_1(Age) + \beta_2(Sex)$$

We'll use 'family = binomial' to run binomial logistic regression...otherwise, this looks very similar to other types of regression we've already done. 

a. Read in the DonnerTable.csv file as DonnerTable

```{r donner_data}
# What effect did age and sex have on the survival rate of the donner party?

# read csv
donnertable = read_csv("DonnerTable.csv")
# outcome of 1 = survival; 0 = death D:
# 0 = yes, 1 = no????

```

b. Binomial logistic regression

```{r blr}

# family = binomial... depends on equation wanted!
# in log odds
donnerblr = glm(Outcome ~ Sex + Age, family = "binomial", data = donnertable)
summary(donnerblr)
# shows that males and higher age had worse survival rate, but know tha tthis is WRONG!!
```

c. Use the model to answer: What are the log odds of survival for a 5 year-old female? The probability of survival?

```{r predict_f_5}

# 1. Create a data frame with variables Sex and Age, containing data "Female" and 5, respectively: 

f5 = data.frame(Sex = "Female", Age = 5)

# 2. Find the log odds of survival for the new data (5 year old female) using predict() function with type = "link":

f5_logodds = predict(donnerblr, newdata = f5, type = "link")
f5_logodds #1.444...

# 3. Exponentiate the log odds to find ODDS of survival for a 5 year old female:

f5_odds = exp(f5_logodds)
f5_odds # the odds of survival for a 5 year old female is 4:1... thus probabilty is 80%.

# Ask: Could we manually find the probability of survival for a 5-year old female? recall: p/(1-p) = ODDS

# 4. Actually, let's just use type = "response" in the predict function, which converts to a probability for us:

f5_prob = predict(donnerblr, newdata = f5, type = "response")
f5_prob # 80.90%

```

d. What is the probability of survival for a 25 year-old male? 

```{r predict_m_25}

# Similarly:

m_25 <- data.frame(Sex = "Male", Age = 25) # Make a new data frame

m_25_prob <- predict(donnerblr, newdata = m_25, type = "response") # Find probability of survival
m_25_prob # only 42.67% survival rate

```

e. Create new sequences of data so that we can graph probabilities for the entire spectrum of ages, designated by sex.

```{r new_data}

seq_age <- rep(seq(from = 0, to = 100), 2) # Create a sequence from 0 to 100, twice (one will be "Male" and one will be "Female")

f_101 <- rep("Female", 101) # Repeat 'Female' 101 times (to match years data)
m_101 <- rep("Male", 101) # Repeat 'Male' 101 times
mf_101 <- c(f_101, m_101) # Combine them into a single vector

# Combine the age and sex sequences into a single data frame - that will be the new data that we have our model make predictions for

donner_newdata <- data.frame(seq_age, mf_101) # MUST make column names match variables in the model!
colnames(donner_newdata) <- c("Age","Sex")

```

f. Now that we have new data to put into our model to have it make predictions, let's go ahead and actually find the predicted probabilities for each Age/Sex combination.

```{r donner_predict}

# Find probabilities using predict (with type = "response"). Include SE (error bars)

predicted_probs = predict(donnerblr, newdata = donner_newdata, type = "response", se.fit = TRUE)

# Coerce outcome into data frame. 

graphdata = data.frame(donner_newdata, predicted_probs$fit, predicted_probs$se.fit)
colnames(graphdata) = c("Age", "Sex", "Probability", "SE")

```

g. Graph results.

```{r donner_graph}

ggplot(graphdata, aes(x = Age, y = Probability)) +
  geom_line(aes(color = Sex)) + 
  geom_ribbon(aes(ymin= Probability - SE, ymax = Probability + SE, fill = Sex, alpha = 0.3))
  
```


# PART 3. Solar irradiation at the 3 locations in 2010 (pull in all together, do some lubridate stuff, etc.)

a. Read in multiple solar irradiation files (for SB (CA), Hilo (HI), and Fairbanks (AK)):
```{r solar_data}

# adding csv files all together :O match anything that begins with solar_radiation
sifull = list.files(pattern = "solar_irradiation_*") %>% 
  map_df(~read_csv(.)) %>% # "." means the line before
  clean_names() #cleaning column names

```

b. Wrangle the data
```{r wrangle_solar}

solartidy = sifull %>% 
  rename(sol_rad = etr_wh_m_2, date = yyyy_mm_dd, time = hh_mm_lst) %>% # renaming columns
  filter(time != "NA") %>% # taking NA values out of time coolumn
  mutate(site = fct_relevel(site, "Hawaii", "Santa Barbara", "Alaska"))

```

c. Use lubridate() functions to convert to times/dates
```{r date_time}

# take this column, reorganize dates as mdy, and then put it back into the column
solartidy$date = mdy(solartidy$date)
solartidy$time = hms(solartidy$time)

```

d. Make an awesome figure of solar irradiation (heat/tile map)
```{r solar_graph}

solargg = ggplot(solartidy, aes(x = date, y = time)) +
  geom_tile(aes(fill = sol_rad)) + 
  scale_fill_gradientn(colors = c("royalblue2", "mediumorchid1", "orange", "yellow")) +  # endless amount of colors to add to maunally made gradient scale
  scale_y_time() +
  facet_wrap(~site, ncol = 1)
solargg

```


#PART 4. gganimate example: total aquaculture production (metric tons) for United States, Brazil, Chile, and Ecuador

a. Get the data, and do some wrangling:
```{r aqua_data}

aq_df = read_csv("aq_wb.csv")

aq_tidy = aq_df %>% 
  filter(country == "Brazil" | country == "Chile" | country == "Ecuador" | country == "United States") %>% 
  gather(year, aq_prod, `1960`:`2016`) %>% 
  filter(year >= 1990) %>% 
  mutate(aq_mil = aq_prod/1000000) %>% 
  select(country, year, aq_mil)
```


b. Read in the fish.png as 'fish'

```{r fish_image}

fish = "fish.png"

```

c. Make a graph...with gganimate!
```{r gganimate_fish}

aqplot = ggplot(aq_tidy, aes(x = as.numeric(year), y = aq_mil, group = country)) +
  geom_line(aes(color = country)) +
  geom_image(aes(image = fish)) + 
  geom_text(aes(label = country, color = country), position = position_nudge(y = 0.04, x = 1), size = 5)
 # + transition_reveal(country, as.numeric(year)) # need to have gganimate (can't install this)

aqplot

```

##END LAB